**Machine learning identifies unique taxa differentiating proximal and distal

### 2017-06-08 submission/manuscript.Rmd
methods added

#### Sample collection
At a baseline visit, subjects were consented and given a home collection stool kit (Source of kit supplies). At least one week prior to the scheduled colonoscopy, subjects were to collect whole stool at home and two swabs of a Fecal Immunohistochemical Test cartridge (Polymedco) and ship the samples to the University. Notably, subjects did not undergo any bowel preparation method prior to sampling. On procedure day, subjects reported to the Michigan Clinical Research Unit at the University of Michigan Health System. Patients were consciously sedated using Fentanyl, Versed and/or Benadryl as appropriate. A flexible sigmoidoscope was first inserted about 25 cm into the colon and endoscopy brush used to collect lumenal/stool contents. Two lumenal samples were collected and the contents immediately deposited into RNAlater (source) and flash-frozen in liquid nitrogen. The brushes were withdrawn and biopsy forceps were used to collect mucosal biopsies on sections of the colon that were pink and free of stool matter. Three (double check this) mucosal biopsies were collected and flash-frozen in RNAlater. These samples comprised the distal or left colon samples. The sigmoidoscope was then withdrawn and a pediatric colonoscope was inserted to reach the ascending colon. Samples were then collected as in the distal colon and the colonoscope withdrawn. All samples were stored at -80 C until study completion.

#### Sample processing and sequencing

>>>>

#### Sample collection
At a baseline visit, subjects were consented and given a home collection stool kit (Source of kit supplies). At least one week prior to the scheduled colonoscopy, subjects were to collect whole stool at home and two swabs of a Fecal Immunohistochemical Test cartridge (Polymedco Inc.) and ship the samples to the University. Notably, subjects did not undergo any bowel preparation method prior to sampling. On procedure day, subjects reported to the Michigan Clinical Research Unit at the University of Michigan Health System. Patients were consciously sedated using Fentanyl, Versed and/or Benadryl as appropriate. A flexible sigmoidoscope was first inserted about 25 cm into the colon and endoscopy brush used to collect lumenal/stool contents. Two lumenal samples were collected and the contents immediately deposited into RNAlater (source) and flash-frozen in liquid nitrogen. The brushes were withdrawn and biopsy forceps were used to collect mucosal biopsies on sections of the colon that were pink and free of stool matter. Three (double check this) mucosal biopsies were collected and flash-frozen in RNAlater. These samples comprised the distal or left colon samples. The sigmoidoscope was then withdrawn and a pediatric colonoscope was inserted to reach the ascending colon. Samples were then collected as in the distal colon and the colonoscope withdrawn. All samples were stored at -80 C until study completion.

#### Sample processing and sequencing
DNA extraction was performed using the PowerMicrobiome DNA/RNA Isolation Kit (MO BIO Laboratories). For tissue biopsies, Bond-Breaker TCEP solution (Fisher) and 2.8mm ceramic beads (MO BIO Laboratories) were added to the bead beating step to enhance DNA recovery from mucosal samples. The resulting DNA was used as template for amplification of the V4 region of the 16S rRNA gene and fragments were sequenced on an Illumina MiSeq as previously described (@Kozich 2013). Sequences were curated and clustered into operational taxonomic units (OTUs) using mothur software as described previously (Schloss 2009). The resultant OTUs assigned taxonomic classification using a naive Bayesian classifier trained using a 16S rRNA gene training set from the Ribosomal Database Project (RDP) (wang 2007).
